### 2021-03-02 02_abstract/abstract_english.tex
Added first part of English abstract and cleaned tex files

TBA.

>>>>

A precise quantification of climate change is crucial in order to assess
optimal mitigation and adaptation strategies. Valuable tools in this context
are \acp{ESM}, which are state-of-the-art climate models that allow numerical
simulations of the complex physical, biological and chemical processes of the
Earth system. Of particular relevance are projections of the future climate,
which enable an assessment of future climate change. Due to the chaotic nature
of the climate system, unknowns in future emission pathways and uncertainties
in the climate models, projections of the future climate are associated with
large uncertainties. In this thesis, we analyze future climate projections from
\acp{ESM} participating in the \ac{CMIP} and reduce corresponding uncertainties
with observations.

### 2021-03-02 ch01_introduction/introduction.tex
Added first part of English abstract and cleaned tex files

Simulations that extrapolate state of the climate system into the future are
called \emph{projections} of the future climate. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on a sustainable development). In many variables that
are relevant for climate change, multi-model projections from \ac{CMIP} show a
large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is the climate sensitivity, which refers to
the change in the \ac{GSAT} that results from a change in the radiative
forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration and the \emph{\ac{TCR}}, which describes the transient
response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
have been assessed with high ranges of $\rangeunit{1.5}{4.5}{K}$ and
$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
\autocite{Stocker2013}. The corresponding inter-model ranges from the
\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
a careful statistical evaluation and further refinement of the output of
multi-model climate projections is necessary in order to reduce associated
uncertainties. State-of-the-art techniques for this are so-called
\emph{emergent constraints}, which use a physically-based inter-model
relationship between an observable quantity of the Earth system and a target
variable to reduce uncertainties in the target variable with observations
\autocite{Allen2002}. An alternative approach is climate model weighting based
on performance (\ie{} the distance of the climate models relative to
observational products) and interdependence (\ie{} the distance between the
climate models among each other) \autocite{Knutti2017a}. These techniques form
the baseline for the new analyses and results presented in this thesis.

>>>>

Simulations that extrapolate state of the climate system into the future are
called \emph{projections} of the future climate. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on a sustainable development). In many variables that
are relevant for climate change, multi-model projections from \ac{CMIP} show a
large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is the climate sensitivity, which refers to
the change in the \ac{GSAT} that results from a change in the radiative
forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration and the \emph{\ac{TCR}}, which describes the transient
response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
have been assessed with high ranges of $\rangeunit{1.5}{4.5}{K}$ and
$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
\autocite{Stocker2013}. The corresponding inter-model ranges from the
\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
a careful statistical evaluation and further refinement of the output of
multi-model climate projections is necessary in order to reduce associated
uncertainties. A state-of-the-art technique for this is the \emph{emergent
  constraints} method, which uses a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. An alternative approach is the weighting of climate
models based on their performance (\ie{} the distance of one model to
observational products) and interdependence (\ie{} the distance of one model to
other climate models) \autocite{Knutti2017a}. These techniques form the
baseline for the new analyses and results presented in this thesis, which
partly utilize methods from a new emerging research field in climate sciences:
\ac{AI} and \ac{ML}.

### 2021-03-02 ch02_scientific_background/scientific_background.tex
Added first part of English abstract and cleaned tex files

Thus, this removal of \ac{CO2} from the atmosphere actively slows down global
warming. However, whether this benefit will persist in the future remains
unclear, which is primarily linked to two feedback processes connecting the
physical climate system and the global carbon cycle: the
\emph{concentration-carbon feedback} and the \emph{climate-carbon feedback}
\autocite{Friedlingstein2006, Gregory2009, Collins2013}. For the terrestrial
biosphere, the concentration-carbon feedback is connected to the \emph{\ac{CO2}
  fertilization effect} \autocite{Walker2020}, that causes an increase of
photosynthesis rates when the atmospheric \ac{CO2} concentration increases,
which in turns removes \ac{CO2} from the atmosphere, forming a negative
feedback. For the ocean, the concentration-carbon feedback is negative as well.
In this case, an elevated atmospheric \ac{CO2} concentration causes an
increased dissolution of \ac{CO2} into the sea, which increases the ocean
carbon uptake. On the other hand, the climate-carbon feedback is thought to be
positive for both the terrestrial biosphere and the ocean
\autocite{Gregory2009}. In the first case, temperature and precipitation
changes due to anthropogenic activities decrease the land carbon uptake because
of increased temperature and water stress on photosynthesis and higher
ecosystem respiration costs, which accelerates global warming due to more
\ac{CO2} that remains in the atmosphere. For the ocean, increased temperatures
lead to a reduction of vertical transport in the ocean resulting from increased
stability and reduced solubility of \ac{CO2} in the sea, which reduces the
ocean carbon uptake and enhances climate change \autocite{Gregory2009}.

@@

The definition of the weights according to \cref{eq:02:knutti_weights} is based
on reasonable and comprehensible principles. However, the exact form of the
equation is purely subjective. Moreover, the additional freedom in choosing a
suitable metric and optimal values for the parameters $\sigma_D$ and $\sigma_S$
adds another level of subjectivity to the weighting scheme, which can partly be
addressed with the introduced leave-one-model-out \ac{CV} setup. Nevertheless,
due to its flexibility, the climate model weighting scheme of
\textcite{Knutti2017a} has already been used for various target variables:
Arctic sea ice \autocite{Knutti2017a}, Antarctic ozone concentrations
\autocite{Amos2020}, North American maximum temperature \autocite{Lorenz2018},
European temperature and precipitation \autocite{Brunner2019, Merrifield2020}
and global warming over the \nth{21} century \autocite{Brunner2020,
  Liang2020a}.

>>>>

Thus, this removal of \ac{CO2} from the atmosphere actively slows down global
warming. However, whether this benefit will persist in the future remains
unclear, which is primarily linked to two feedback processes connecting the
physical climate system and the global carbon cycle: the
\emph{concentration-carbon feedback} and the \emph{climate-carbon feedback}
\autocite{Friedlingstein2006, Gregory2009, Collins2013}. For the terrestrial
biosphere, the concentration-carbon feedback is connected to the \emph{\ac{CO2}
  fertilization effect} \autocite{Walker2020}, that causes an increase of
photosynthesis rates when the atmospheric \ac{CO2} concentration increases,
which in turns removes \ac{CO2} from the atmosphere, forming a negative
feedback. For the ocean, the concentration-carbon feedback is negative as well.
In this case, an elevated atmospheric \ac{CO2} concentration causes an
increased dissolution of \ac{CO2} into the sea, which increases the ocean
carbon uptake. On the other hand, the climate-carbon feedback is thought to be
positive for both the terrestrial biosphere and the ocean
\autocite{Gregory2009}. In the first case, temperature and \acl{PR} changes due
to anthropogenic activities decrease the land carbon uptake because of
increased temperature and water stress on photosynthesis and higher ecosystem
respiration costs, which accelerates global warming due to more \ac{CO2} that
remains in the atmosphere. For the ocean, increased temperatures lead to a
reduction of vertical transport in the ocean resulting from increased stability
and reduced solubility of \ac{CO2} in the sea, which reduces the ocean carbon
uptake and enhances climate change \autocite{Gregory2009}.

@@

The definition of the weights according to \cref{eq:02:knutti_weights} is based
on reasonable and comprehensible principles. However, the exact form of the
equation is purely subjective. Moreover, the additional freedom in choosing a
suitable metric and optimal values for the parameters $\sigma_D$ and $\sigma_S$
adds another level of subjectivity to the weighting scheme, which can partly be
addressed with the introduced leave-one-model-out \ac{CV} setup. Nevertheless,
due to its flexibility, the climate model weighting scheme of
\textcite{Knutti2017a} has already been used for various target variables:
Arctic sea ice \autocite{Knutti2017a}, Antarctic ozone concentrations
\autocite{Amos2020}, North American maximum temperature \autocite{Lorenz2018},
European temperature and \acl{PR} \autocite{Brunner2019, Merrifield2020} and
global warming over the \nth{21} century \autocite{Brunner2020, Liang2020a}.

### 2021-03-02 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added first part of English abstract and cleaned tex files

Various improvements of the underlying physical, chemical and biological
processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
represent the coupled Earth system in more detail. Since many of these
processes directly influence the models' \ac{ECS} \autocite{Forster2020},
determining reasons for the increased \ac{ECS} in the \acs{CMIP}6 ensemble is
highly non-trivial. As discussed in detail by \textcite{Meehl2020}, one
possible reason is the consideration of aerosol-cloud interactions. New
prognostic aerosol schemes added to some \acs{CMIP}6 models that include
aerosol-cloud interactions could have produced overly large negative radiative
forcing, which then required a stronger model response to \ac{GHG} forcing in
order to reproduce the observed historical temperature trend. In fact,
\textcite{Meehl2020} find a weak inter-model relationship between the aerosol
forcing and \ac{ECS} for some \acs{CMIP}6 models, with larger negative
present-day aerosol forcing associated with higher \ac{ECS}. Such relationships
have also been found for previous climate model generations
\autocite{Kiehl2007, Forster2013}. However, due to the varying aerosol forcing
over time, these relations cannot be used to tune \ac{ECS} based on the models'
responses to aerosols over the different periods of the \nth{20} century
\autocite{Dittus2020}.

>>>>

Various improvements of the underlying physical, biological and chemical
processes have been introduced to the \acs{CMIP}6 \acp{ESM} in order to
represent the coupled Earth system in more detail. Since many of these
processes directly influence the models' \ac{ECS} \autocite{Forster2020},
determining reasons for the increased \ac{ECS} in the \acs{CMIP}6 ensemble is
highly non-trivial. As discussed in detail by \textcite{Meehl2020}, one
possible reason is the consideration of aerosol-cloud interactions. New
prognostic aerosol schemes added to some \acs{CMIP}6 models that include
aerosol-cloud interactions could have produced overly large negative radiative
forcing, which then required a stronger model response to \ac{GHG} forcing in
order to reproduce the observed historical temperature trend. In fact,
\textcite{Meehl2020} find a weak inter-model relationship between the aerosol
forcing and \ac{ECS} for some \acs{CMIP}6 models, with larger negative
present-day aerosol forcing associated with higher \ac{ECS}. Such relationships
have also been found for previous climate model generations
\autocite{Kiehl2007, Forster2013}. However, due to the varying aerosol forcing
over time, these relations cannot be used to tune \ac{ECS} based on the models'
responses to aerosols over the different periods of the \nth{20} century
\autocite{Dittus2020}.

### 2021-03-02 ch05_paper_ecs/paper_ecs.tex
Added overall summary
Added first part of English abstract and cleaned tex files

The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
previous \namecref{ch:04:papers_ecs_tcr_assessment} begs the question about the
interpretation and credibility of these results. Since the \acs{CMIP}6 models
provide crucial input for the upcoming international climate assessment of the
\ac{IPCC} (\acs{AR}6), this is not only important for the scientific community,
but also directly affects society and policymakers. Due to the many new and
improved processes included to the models, the \acs{CMIP}6 ensemble is arguably
more realistic than its predecessor generations. However, does this
automatically make results derived from \acs{CMIP}6 models more trustworthy
than those derived from earlier \ac{CMIP} models? In general, the answer to
this question is \enquote{no}. Even the \acs{CMIP}6 models are far from fully
simulating all aspects and processes of the full vastly complex Earth system,
which could introduce new biases which have not been present in earlier
\ac{CMIP} models. For example, the apparent increase in climate sensitivity in
the \acs{CMIP}6 models might be wrongly caused by missing processes that cancel
out the effects of the newly included processes. Thus, to assess the
credibility of \ac{ESM} results it is vital to take other sources of
information into account.

@@

\textcite{Tian2015} finds a link between mid-tropospheric humidity over the
tropical Pacific and simulated moisture, precipitation, clouds, large-scale
circulation, and thus \ac{ECS} in \acs{CMIP}3 and \acs{CMIP}5 models. The study
explains this link with the similarity of mid-tropospheric humidity and
precipitation patterns as both are related to the \ac{ITCZ}. The proposed
tropical mid-tropospheric humidity asymmetry index to constrain \ac{ECS} is
defined as relative bias (in $\si{\%}$) in simulated annual mean $500
\unit{hPa}$ specific humidity averaged over the Southern Hemisphere tropical
Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$, $\range{80
  \unit{\degree W}}{120 \unit{\degree E}}$) minus the bias averaged over the
Northern Hemisphere tropical Pacific ($\rangeunit{0}{20}{\degree N}$,
$\range{80 \unit{\degree W}}{120 \unit{\degree E}}$) when compared with
observations. Similar to the SU constraint, the index proposed by
\textcite{Tian2015} seems to be related to the emergent constraint by
\textcite{Fasullo2012}, who find correlations between relative humidity of the
middle and upper troposphere and \ac{ECS}. Here, we use humidity observations
from AIRS \autocite{Aumann2003} over the time period \range{2003}{2005} as the
reference dataset. We assess a $66 \unit{\%}$ \ac{ECS} range of
$\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5 ($R^2 = 0.24$) and
$\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 = 0.06$).
\textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
significance of the emergent relationship drops massively from $p = 0.0089$ in
\acs{CMIP}5 to $p = 0.1348$ in \acs{CMIP}6.

@@

In addition to the humidity index, \textcite{Tian2015} proposes an emergent
constraint on \ac{ECS} based on the southern \ac{ITCZ} index
\autocite{Bellucci2010, Hirota2011}. This index is defined as the
climatological annual mean precipitation bias averaged over the southeastern
Pacific ($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$,
$\rangeunit{150}{100}{\degree W}$). The southern \ac{ITCZ} index is calculated
in $\si{\mmday{}}$ and dominated by the so-called double-\ac{ITCZ}, a common
problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds a link
between double-\ac{ITCZ} bias and simulated moisture, precipitation, clouds,
and large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He argues
that this could explain the link between the double-\acs{ITCZ} bias and
\ac{ECS}. As reference data, we use observed precipitation data for the years
\range{1986}{2005} from GPCP \autocite{Adler2003}. We calculate an \ac{ECS} $66
\unit{\%}$ confidence range of $\pmrangeunit{3.87}{0.67}{K}$ for \acs{CMIP}5
($R^2 = 0.33$) and $\pmrangeunit{3.84}{1.09}{K}$ for \acs{CMIP}6 ($R^2 <
0.01$). \textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
emergent relationship shows a much lower statistical significance in
\acs{CMIP}6 ($p = 0.8236$) than in \acs{CMIP}5 ($p = 0.0013$).

>>>>

The massive spread in \ac{ECS} within the \acs{CMIP}6 ensemble discussed in the
previous \namecref{ch:04:papers_ecs_tcr_assessment} begs the question about the
interpretation and credibility of these results. Since the \acs{CMIP}6 models
provide crucial input for the upcoming international climate assessment of the
\ac{IPCC} (\acs{AR}6), this is not only important for the scientific community,
but also directly affects policymakers and society. Due to the many new and
improved processes included to the models, the \acs{CMIP}6 ensemble is arguably
more realistic than its predecessor generations. However, does this
automatically make results derived from \acs{CMIP}6 models more trustworthy
than those derived from earlier \ac{CMIP} models? In general, the answer to
this question is \enquote{no}. Even the \acs{CMIP}6 models are far from fully
simulating all aspects and processes of the full vastly complex Earth system,
which could introduce new biases which have not been present in earlier
\ac{CMIP} models. For example, the apparent increase in climate sensitivity in
the \acs{CMIP}6 models might be wrongly caused by missing processes that cancel
out the effects of the newly included processes. Thus, to assess the
credibility of \ac{ESM} results it is vital to take other sources of
information into account.

@@

\textcite{Tian2015} finds a link between mid-tropospheric humidity over the
tropical Pacific and simulated moisture, \acl{PR}, clouds, large-scale
circulation, and thus \ac{ECS} in \acs{CMIP}3 and \acs{CMIP}5 models. The study
explains this link with the similarity of mid-tropospheric humidity and
\acl{PR} patterns as both are related to the \ac{ITCZ}. The proposed tropical
mid-tropospheric humidity asymmetry index to constrain \ac{ECS} is defined as
relative bias (in $\si{\%}$) in simulated annual mean $500 \unit{hPa}$ specific
humidity averaged over the Southern Hemisphere tropical Pacific ($\range{30
  \unit{\degree S}}{0 \unit{\degree N}}$, $\range{80 \unit{\degree W}}{120
  \unit{\degree E}}$) minus the bias averaged over the Northern Hemisphere
tropical Pacific ($\rangeunit{0}{20}{\degree N}$, $\range{80 \unit{\degree
    W}}{120 \unit{\degree E}}$) when compared with observations. Similar to
the SU constraint, the index proposed by \textcite{Tian2015} seems to be
related to the emergent constraint by \textcite{Fasullo2012}, who find
correlations between relative humidity of the middle and upper troposphere and
\ac{ECS}. Here, we use humidity observations from AIRS \autocite{Aumann2003}
over the time period \range{2003}{2005} as the reference dataset. We assess a
$66 \unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.88}{0.75}{K}$ for \acs{CMIP}5
($R^2 = 0.24$) and $\pmrangeunit{4.15}{1.10}{K}$ for \acs{CMIP}6 ($R^2 =
0.06$). \textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The
significance of the emergent relationship drops massively from $p = 0.0089$ in
\acs{CMIP}5 to $p = 0.1348$ in \acs{CMIP}6.

@@

In addition to the humidity index, \textcite{Tian2015} proposes an emergent
constraint on \ac{ECS} based on the southern \ac{ITCZ} index
\autocite{Bellucci2010, Hirota2011}. This index is defined as the
climatological annual mean \acl{PR} bias averaged over the southeastern Pacific
($\range{30 \unit{\degree S}}{0 \unit{\degree N}}$,
$\rangeunit{150}{100}{\degree W}$). The southern \ac{ITCZ} index is calculated
in $\si{\mmday{}}$ and dominated by the so-called double-\ac{ITCZ}, a common
problem in many \acs{CMIP}5 climate models. \textcite{Tian2015} finds a link
between double-\ac{ITCZ} bias and simulated moisture, \acl{PR}, clouds, and
large-scale circulation in \acs{CMIP}3 and \acs{CMIP}5 models. He argues that
this could explain the link between the double-\acs{ITCZ} bias and \ac{ECS}. As
reference data, we use observed \acl{PR} data for the years \range{1986}{2005}
from GPCP \autocite{Adler2003}. We calculate an \ac{ECS} $66 \unit{\%}$
confidence range of $\pmrangeunit{3.87}{0.67}{K}$ for \acs{CMIP}5 ($R^2 =
0.33$) and $\pmrangeunit{3.84}{1.09}{K}$ for \acs{CMIP}6 ($R^2 < 0.01$).
\textcite{Tian2015} specifies a best estimate of $4.0 \unit{K}$. The emergent
relationship shows a much lower statistical significance in \acs{CMIP}6 ($p =
0.8236$) than in \acs{CMIP}5 ($p = 0.0013$).

### 2021-03-02 end_acronyms/acronyms.tex
Added figures of appendix B to tex file
Added text to chapter 06
Finished chapter 07
Finished chapter 01
Added first part of English abstract and cleaned tex files

\begingroup
\setstretch{0.5}
\begin{acronym}
  \acro{1BGC}[1\%BGC]{biogeochemically-coupled simulation where the %
    atmospheric CO\textsubscript{2} concentration increases by $1 \unit{\%}$ %
    per year}
  \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
  \acro{AOGCM}{atmosphere-ocean general circulation model}
  \acro{AR}{Assessment Report}
  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
  \acro{CESM}{Community Earth System Model}
  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
  \acro{CMIP}{Coupled Model Intercomparison Project}
  \acro{CH4}[CH\textsubscript{4}]{methane}
  \acro{CMOR}{Climate Model Output Rewriter}
  \acro{CO}{carbon monoxide}
  \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
  \acro{CRE}{cloud radiative effect}
  \acro{CV}{cross-validation}
  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
  \acro{DJF}{December-January-February}
  \acro{DLR}{Deutsches Zentrum für Luft- und Raumfahrt (German Aerospace %
    Center)}
  \acro{ECS}{effective climate sensitivity}
  \acro{ENSO}{El Niño-Southern Oscillation}
  \acro{ESGF}{Earth System Grid Federation}
  \acro{ESM}{Earth system model}
  \acro{ESMValCore}{core functionalities of the ESMValTool}
  \acro{ESMValTool}{Earth System Model Evaluation Tool}
  \acro{GBRT}{gradient boosted regression tree}
  \acro{GHG}{greenhouse gas}
  \acro{GPP}{gross primary production}
  \acro{GSAT}{global mean near-surface air temperature}
  \acro{GtC}{gigatonnes of carbon}
  \acro{IAM}{integrated assessment model}
  \acro{IPA}{Institut für Physik der Atmosphäre (Institute of Atmospheric %
    Physics)}
  \acro{IPCC}{Intergovernmental Panel on Climate Change}
  \acro{ITCZ}{Intertropical Convergence Zone}
  \acro{JJA}{June-July-August}
  \acro{KUM}{Cape Kumukahi, Hawaii}
  \acro{LAI}{leaf area index}
  \acro{LASSO}{least absolute shrinkage and selection operator}
  \acro{LIME}{local interpretable model-agnostic explanations}
  \acro{LongRunMIP}{Long Run Model Intercomparison Project}
  \acro{LTMI}{lower-tropospheric mixing index}
  \acro{MBLC}{marine boundary layer cloud}
  \acro{MDER}{multiple diagnostic ensemble regression}
  \acro{MIP}{Model Intercomparison Project}
  \acro{ML}{machine learning}
  \acro{MMM}{multi-model mean}
  \acro{NAO}{North Atlantic Oscillation}
  \acro{NCAR}{National Center for Atmospheric Research}
  \acro{netCDF}{Network Common Data Form}
  \acro{PDF}{probability density function}
  \acro{ppm}{parts per million}
  \acro{PR}{precipitation}
  \acro{RCP}{Representative Concentration Pathway}
  \acro{RMSE}{root-mean-square error}
  \acro{RMSEP}{root-mean-square error of prediction}
  \acro{RSDS}{downwelling solar radiation at surface}
  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
  \acro{SSP}{Shared Socioeconomic Pathway}
  \acro{SST}{sea surface temperature}
  \acro{T}{near-surface air temperature}
  \acro{TCR}{transient climate response}
  \acro{TLC}{tropical low-level cloud}
  \acro{TOA}{top of the atmosphere}
  \acro{WCRP}{World Climate Research Programme}
  \acro{WGCM}{Working Group on Coupled Modelling}
\end{acronym}
\endgroup

>>>>

\begingroup
\setstretch{0.5}
\begin{acronym}
  \acro{1BGC}[1\%BGC]{biogeochemically-coupled simulation where the %
    atmospheric CO\textsubscript{2} concentration increases by $1 \unit{\%}$ %
    per year}
  \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
  \acro{AI}{artificial intelligence}
  \acro{AOGCM}{atmosphere-ocean general circulation model}
  \acro{AR}{Assessment Report}
  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
  \acro{CESM}{Community Earth System Model}
  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
  \acro{CMIP}{Coupled Model Intercomparison Project}
  \acro{CH4}[CH\textsubscript{4}]{methane}
  \acro{CMOR}{Climate Model Output Rewriter}
  \acro{CO}{carbon monoxide}
  \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
  \acro{CRE}{cloud radiative effect}
  \acro{CV}{cross-validation}
  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
  \acro{DJF}{December-January-February}
  \acro{ECS}{effective climate sensitivity}
  \acro{ENSO}{El Niño-Southern Oscillation}
  \acro{ESGF}{Earth System Grid Federation}
  \acro{ESM}{Earth system model}
  \acro{ESMValCore}{core functionalities of the ESMValTool}
  \acro{ESMValTool}{Earth System Model Evaluation Tool}
  \acro{GBRT}{gradient boosted regression tree}
  \acro{GHG}{greenhouse gas}
  \acrodefplural{GHG}{greenhouse gases}
  \acro{GPP}{gross primary production}
  \acro{GSAT}{global mean near-surface air temperature}
  \acro{GtC}{gigatonnes of carbon}
  \acro{IAM}{integrated assessment model}
  \acro{IPCC}{Intergovernmental Panel on Climate Change}
  \acro{ITCZ}{Intertropical Convergence Zone}
  \acro{JJA}{June-July-August}
  \acro{KUM}{Cape Kumukahi, Hawaii}
  \acro{LAI}{leaf area index}
  \acro{LASSO}{least absolute shrinkage and selection operator}
  \acro{LIME}{local interpretable model-agnostic explanations}
  \acro{LongRunMIP}{Long Run Model Intercomparison Project}
  \acro{LTMI}{lower-tropospheric mixing index}
  \acro{MBLC}{marine boundary layer cloud}
  \acro{MDER}{multiple diagnostic ensemble regression}
  \acro{MIP}{Model Intercomparison Project}
  \acro{ML}{machine learning}
  \acro{MMM}{multi-model mean}
  \acro{N2O}[N\textsubscript{2}O]{nitrous oxide}
  \acro{NAO}{North Atlantic Oscillation}
  \acro{NCAR}{National Center for Atmospheric Research}
  \acro{netCDF}{Network Common Data Form}
  \acro{PDF}{probability density function}
  \acro{ppm}{parts per million}
  \acro{PR}{precipitation}
  \acro{RCP}{Representative Concentration Pathway}
  \acro{RMSE}{root mean square error}
  \acro{RMSEP}{root mean square error of prediction}
  \acro{RSDS}{incoming solar radiation at surface}
  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
  \acro{SPE}{standard prediction error}
  \acro{SSP}{Shared Socioeconomic Pathway}
  \acro{SST}{sea surface temperature}
  \acro{T}{near-surface air temperature}
  \acro{TCR}{transient climate response}
  \acro{TLC}{tropical low-level cloud}
  \acro{TOA}{top of the atmosphere}
  \acro{WCRP}{World Climate Research Programme}
  \acro{WGCM}{Working Group on Coupled Modelling}
\end{acronym}
\endgroup

### 2021-03-02 end_appendix/appendix.tex
Added overall summary
Added first part of English abstract and cleaned tex files

\chapter{Appendix}

@@

\begin{table}[!h]
  \centering
  \begin{tabulary}{\columnwidth}{>{\em}L L}
    \toprule
    \ignorecolumntype{l}{Variable short name} & Description \\
    \midrule
    cl & Cloud area fraction \\
    clt & Total cloud area fraction \\
    hur & Relative humidity \\
    hus & Specific humidity \\
    pr & Precipitation \\
    rsdt & \Acf{TOA} incoming shortwave radiation \\
    rsut & \acs{TOA} outgoing shortwave radiation \\
    rsutcs & Clear-sky \acs{TOA} outgoing shortwave radiation \\
    ta & Air temperature \\
    tas & Surface air temperature \\
    tos & Sea surface temperature \\
    ts & Surface temperature \\
    va & Northward wind speed \\
    wap & Vertical velocity \\
    \bottomrule
  \end{tabulary}
  \caption{Overview of the variables used in \cref{ch:05:paper_ecs}
    (\nameref*{ch:05:paper_ecs}). More details are given in
    \cref{tab:05:overview_emergent_constraints}, which lists the variables
    used for each emergent constraint.}
  \label{tab:app:a:overview_variables}
\end{table}

>>>>

\addchap{Appendix}

@@

\begin{table}[!h]
  \centering
  \begin{tabulary}{\columnwidth}{>{\em}L L}
    \toprule
    \ignorecolumntype{l}{Variable short name} & Description \\
    \midrule
    cl & Cloud area fraction \\
    clt & Total cloud area fraction \\
    hur & Relative humidity \\
    hus & Specific humidity \\
    pr & \Acl{PR} \\
    rsdt & \Acf{TOA} incoming shortwave radiation \\
    rsut & \acs{TOA} outgoing shortwave radiation \\
    rsutcs & Clear-sky \acs{TOA} outgoing shortwave radiation \\
    ta & Air temperature \\
    tas & Surface air temperature \\
    tos & Sea surface temperature \\
    ts & Surface temperature \\
    va & Northward wind speed \\
    wap & Vertical velocity \\
    \bottomrule
  \end{tabulary}
  \caption{Overview of the variables used in \cref{ch:05:paper_ecs}
    (\nameref*{ch:05:paper_ecs}). More details are given in
    \cref{tab:05:overview_emergent_constraints}, which lists the variables
    used for each emergent constraint.}
  \label{tab:app:a:overview_variables}
\end{table}
